### 2017-06-20 manuscript/manuscript.Rmd
Improve writing. Add note on when to commit.

For example, after working for a while on the project, you could have added two files inside well-organized subfolders of the project. The project would like Figure \@ref(fig:advanced-project) when viewed with a file explorer:

@@

Running `git status` now tells that there are two new files (possible empty folders are ignored). That is, since the previous commit, two files have been added to the project; one a pdf with some administrative information (`ethics-info.pdf`), the other one is an R [@r_core_team_r:_2017] script for a pre-planned power analysis (`plan_n.R`). You would probably like to track any changes to the power analysis script, but the ethics information file isn't really something that you need to keep track of---at least for this example. It would be laborious to constantly keep ignoring it when committing changes to Git (especially if we'd like to commit multiple files simultaneously with `git add .`) Git has an elegant solution to specifying which files to keep track of. Because by default all files are monitored, Git uses a special file for instructing which files are to be *ignored*.

@@

To make Git ignore files, you simply add a plain text file called `.gitignore` to the home folder of the repository. You can use any text editor to create this file. Each row of this file should specify a file or a folder (or a regular expression) that Git should ignore. In the current example you could make Git ignore the `admin/` folder entirely, and any file with the .pdf extension inside `manuscript/`. The example .gitignore file would look like this (lines beginning with `#` are comments):

@@

Re-running `git status` now only shows the `plan_n.R` file (and the newly created .gitignore file, which is also under version control, naturally.)

Because there are now two untracked files that are not in .gitignore (`.gitignore` and `plan_n.R`), and you usually should aim to maintain a clean history for the project, you can create two separate commits: One for the Git ignore file, and one for the power analysis file.

@@

After this last commit, you can at any time come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file when it was first created. For instance, if new information suggests that you should change the assumed effect size in the power analysis, you can simply edit and save the file, then add and commit the changes to Git with a helpful message that logs this important event in Git's history.

@@

[^csv]: `.csv` stands for comma separated values, and is a plain text file commonly used to store two-dimensional data.
[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple. We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
[^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
[^cli-links]: If using a text-based command line seems challenging, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT offers a free online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
[^git-link-submodules]: For advanced users, Git *submodules* allow linking projects to each other, or organizing more complex projects into projects and their sub-projects (<https://git-scm.com/book/en/v2/Git-Tools-Submodules>).
[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git in standard Psychology studies.
[^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>
[^git-branches]: A particularly powerful approach for trying new features is **branching**: The project can be duplicated to a new branch and modified, then merged back to the main branch after work on the new feature is complete---or the new branch can be discarded if the work ended unsatisfactory. Branches are outside the scope of this tutorial, for more information see the Git website (<https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell>).
[^subversion]: <https://subversion.apache.org/>
[^mercurial]: <https://www.mercurial-scm.org/>
[^git-folder]: There are no visible changes to a folder once it is tracked by Git. Once Git is initialized in a folder, the only change is that a hidden folder, called `.git` is added, but users do not need to interact with it directly.
[^plain-text]: Plain text has many advantages over proprietary file formats such as Microsoft Word's .docx files. Briefly, plain text is both human and computer readable (.docx are unreadable without a copy of Microsoft Word), is both forward and backward compatible (there will always, and has always been, software capable of reading it), and plain text takes very little space. The file extension of plain text doesn't matter much, but we recommend using either `.txt` because it is widely recognized, `.md` for markdown syntax, or in the case of a readme file, simply not using any file extension.
[^dot-shortcut]: For the `.` shortcut to work, you must currently be in the project's root directory. If you are not in the root directory, you can use `-A` shortcut instead; however, we recommend that you always ensure that you are in the project's root directory when running any Git commands.

>>>>

For example, after working for a while on the project, you could have added two files inside well-organized subfolders of the project. The project would like Figure \@ref(fig:advanced-project) when viewed with a file explorer.

@@

Running `git status` now tells that there are two new files (possible empty folders are ignored). That is, since the previous commit, two files have been added to the project; one a pdf with some administrative information (`ethics-info.pdf`), the other one is an R [@r_core_team_r:_2017] script for a pre-planned power analysis (`plan_n.R`). You would probably like to track any changes to the power analysis script, but the ethics information file isn't really something that you need to keep track of---at least for this example. It would be laborious to constantly keep ignoring it when committing changes to Git (especially if you'd like to add multiple files simultaneously with `git add .`) Git has an elegant solution to specifying which files to keep track of. Because by default all files are monitored, Git uses a special file for instructing which files are to be *ignored*.

@@

To make Git ignore files, you simply add a plain text file called `.gitignore` to the home folder of the repository. You can use any text editor to create this file. Each row of this file should specify a file or a folder (or a regular expression) that Git should ignore. In the current example you could make Git ignore the `admin/` folder entirely, and any file with the .pdf extension inside `manuscript/`. The example `.gitignore` file would look like this (lines beginning with `#` are comments, and only read by humans):

@@

Re-running `git status` now only shows the `plan_n.R` file (and the newly created `.gitignore` file, which is also under version control, naturally.)

Because there are now two untracked files that are not specified to be ignored in `.gitignore` (`.gitignore` and `plan_n.R`), and you usually should aim to maintain a clean commit history for the project, you can create two separate commits: One for the `.gitignore` file, and one for the power analysis file.[^when-to-commit]

@@

After this last commit you can, at any time in the future, come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file when it was first created. For instance, if new information suggests that you should change the assumed effect size in the power analysis, you can simply edit and save the file, then add and commit the changes to Git with a helpful message that logs this important event in Git's history.

@@

[^csv]: `.csv` stands for comma separated values, and is a plain text file commonly used to store two-dimensional data.
[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple. We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
[^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
[^cli-links]: If using a text-based command line seems challenging, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT offers a free online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
[^git-link-submodules]: For advanced users, Git *submodules* allow linking projects to each other, or organizing more complex projects into projects and their sub-projects (<https://git-scm.com/book/en/v2/Git-Tools-Submodules>).
[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git in standard Psychology studies.
[^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>
[^git-branches]: A particularly powerful approach for trying new features is **branching**: The project can be duplicated to a new branch and modified, then merged back to the main branch after work on the new feature is complete---or the new branch can be discarded if the work ended unsatisfactory. Branches are outside the scope of this tutorial, for more information see the Git website (<https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell>).
[^subversion]: <https://subversion.apache.org/>
[^mercurial]: <https://www.mercurial-scm.org/>
[^git-folder]: There are no visible changes to a folder once it is tracked by Git. Once Git is initialized in a folder, the only change is that a hidden folder, called `.git` is added, but users do not need to interact with it directly.
[^plain-text]: Plain text has many advantages over proprietary file formats such as Microsoft Word's .docx files. Briefly, plain text is both human and computer readable (.docx are unreadable without a copy of Microsoft Word), is both forward and backward compatible (there will always, and has always been, software capable of reading it), and plain text takes very little space. The file extension of plain text doesn't matter much, but we recommend using either `.txt` because it is widely recognized, `.md` for markdown syntax, or in the case of a readme file, simply not using any file extension.
[^dot-shortcut]: For the `.` shortcut to work, you must currently be in the project's root directory. If you are not in the root directory, you can use `-A` shortcut instead; however, we recommend that you always ensure that you are in the project's root directory when running any Git commands.
[^when-to-commit]: It is entirely up to the user to decide what to commit and when. However, it is best practice to commit often while making incremental changes. Each commit should aim to solve one problem, introduce one new idea, or---more generally---do one thing. This way, when the commit history is reviewed later, it is easy to find and come back to a specific change.
